---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# latex2r

The goal of latex2r is to translate latex formulas to R code.

THE PACKAGE IS UNDER VERY EARLY DEVELOPMENT, DO NOT USE!

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tomicapretto/latex2r")
```
## Example


```{r example}
library(latex2r)
latex2r("\\beta_1^{\\frac{x+1}{x^2 \\cdot y}}")
```

For those who dare to go a little further

```{r example2}
eval(parse(text = latex2r::latex2r("\\tanh(x_1 ^ 2)")), envir = list(x_1 = 2))
```

For those who are crazy enough to navigate the deepest waters of R, you can see
the [source code](R/latex2fun.R) of `latex2fun`.

NB: The following is very experimental, but shows the power of what you can do with 
this package (and with R).

```{r example3, dpi=500, out.width="75%", fig.align="center"}
x = seq(-2*pi, 2*pi, length.out = 500)
f = latex2fun("\\sin{a * x}^2 + \\cos{b * x} ^2")
print(f)

y = f(x = x, a = 2, b = 3)
par(mar = c(2, 2, 2, 2))
plot(x, y, type = "l")
```
